Eye Fitting Straight Lines in the Modern Era (JCGS-22-045)

### 2022-08-06 analysis/you-draw-it-eyefitting-analysis.Rmd
reran maunscript knit

**Rmd file containing code can be accessed [here](https://github.com/earobinson95/Eye-Fitting-Straight-Lines-in-the-Modern-Era/tree/main/analysis).**

@@

@@

```{r}

eyefitting_parameter_details <- read_csv("eyefitting_parameter_details.csv")
eyefitting_parameter_details %>% knitr::kable()

@@

# write.csv(eyefitting_sim_data, "eyefitting-simulated-data.csv", row.names = F, na = "")

eyefitting_example_simplot <- eyefitting_sim_data %>%
  filter(dataset == "point_data") %>%
  # filter(dataset %in% c("F", "N", "S") | (x < 16 & x > 4)) %>%
  mutate(parm_id = factor(parm_id, levels = c("S", "F", "V", "N"))) %>%
  dplyr::rename(`Parameter Choice` = parm_id) %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(size = 1) +
  facet_wrap(~`Parameter Choice`, ncol = 4) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
  legend.position = "none",
  plot.title   = element_text(size = 12, hjust = 0),
  axis.text    = element_text(size = 12),
  axis.title   = element_text(size = 12),
  legend.title = element_text(size = 12),
  legend.text  = element_text(size = 12),
  # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
  # strip.background = element_rect(size = 0.5),
  legend.key.size = unit(1, "line")
) +
  scale_y_continuous(breaks = seq(-10, 20, 5))
eyefitting_example_simplot
```

@@

# save data to csv - already run once and pushed to GitHub.
# write.csv(eyefitting_model_data, file = "eyefitting-model-data.csv", row.names = F, na = "")
```

@@

De-identified participant data collected in the study and used for analyses are available to be downloaded from GitHub at [here](https://github.com/earobinson95/Eye-Fitting-Straight-Lines-in-the-Modern-Era/raw/main/analysis/eyefitting-model-data.csv).

```{r data}
eyefitting_model_data  <- read_csv("https://github.com/earobinson95/Eye-Fitting-Straight-Lines-in-the-Modern-Era/raw/main/analysis/eyefitting-model-data.csv")

@@

```{r}
# Example
eyefitting_model_data %>%
  filter(participant_id == "60f510d6896cddc508ea6d3de4b58234", parm_id == "F") %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = yols, color = "OLS", linetype = "OLS")) +
  geom_line(aes(y = ypca, color = "PA", linetype = "PA")) +
  geom_line(aes(y = ydrawn, color = "Drawn", linetype = "Drawn")) +
  geom_point(data = eyefitting_simulated_data %>%
               filter(dataset == "point_data", participant_id == "60f510d6896cddc508ea6d3de4b58234", parm_id == "F"),
             aes(x = x, y = y)) +
  facet_wrap(~ parm_id) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_y_continuous("y") +
  scale_color_manual("", values = c("black", "steelblue", "orange2")) +
  scale_linetype_manual("", values = c("dashed", "solid", "solid"))
```

@@

```{r y-ols-plot}

eyefitting_model_data %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = ydrawn, group = plot_id), alpha = 0.5, color = "steelblue") +
  geom_line(alpha = 0.2, aes(y = yols, group = participant_id)) +
  facet_wrap(~ parm_id) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_x_continuous(limits = c(0, 20)) +
  ggtitle("OLS")
```

@@

```{r y-pca-plot}
eyefitting_model_data %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = ydrawn, group = plot_id), alpha = 0.5, color = "steelblue") +
  geom_line(alpha = 0.2, aes(y = ypca, group = participant_id)) +
  facet_wrap(~ parm_id) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_x_continuous(limits = c(0, 20)) +
  ggtitle("PA")
```

@@

```{r residual-plot}
eyefitting_model_data %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = residual_ols_drawn, group = plot_id, color = "OLS"), alpha = 0.3) +
  geom_line(aes(y = residual_pca_drawn, group = plot_id, color = "PA"), alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~parm_id) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_x_continuous(limits = c(0, 20)) +
  scale_color_manual("Estimate", values = c("steelblue", "orange"))
```

@@

+ parm_ID (S, F, V, N)
+ x (0, 20)

@@

```{r lmer, fig.height = 9}
# OLS
eyefitting.ols.lmer <- lmer(residual_ols_drawn ~ -1 + parm_id + x:parm_id + (1|participant_id),
                       data = eyefitting_model_data)
anova(eyefitting.ols.lmer)

@@

# Plot Predictions
eyefitting.lmer.plot <- eyefitting.preds.lmer %>%
  filter((parm_id %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(parm_id = factor(parm_id, levels = c("S", "F", "V", "N"))) %>%
  dplyr::rename(`Parameter Choice` = parm_id) %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data %>% mutate(parm_id = factor(parm_id, levels = c("S", "F", "V", "N"))) %>% dplyr::rename(`Parameter Choice` = parm_id), 
            aes(x = x, y = residual_ols_drawn, group = plot_id, color = "OLS"), alpha = 0.2) +
  geom_line(data = eyefitting_model_data %>% mutate(parm_id = factor(parm_id, levels = c("S", "F", "V", "N"))) %>% dplyr::rename(`Parameter Choice` = parm_id), 
            aes(x = x, y = residual_pca_drawn, group = plot_id, color = "PCA"), alpha = 0.2) +
  geom_ribbon(aes(ymin = asymp.LCL.ols, ymax = asymp.UCL.ols, fill = "OLS"), color = NA, alpha = 0.4) +
  geom_line(aes(y = emmean.ols, color = "OLS")) +
  geom_ribbon(aes(ymin = asymp.LCL.pca, ymax = asymp.UCL.pca, fill = "PA"), color = NA, alpha = 0.4) +
  geom_line(aes(y = emmean.pca, color = "PCA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, ncol = 1) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "right",
        plot.title   = element_text(size = 12, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line")
        ) +
  scale_y_continuous("Residual") +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("LMER fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PA"))

@@

```{r gamm, fig.height = 9}
# OLS
eyefitting.ols.gamm <- bam(residual_ols_drawn ~ parm_id + s(x, by = parm_id) + 
                             s(participant_id, bs = "re") +
                             s(x,participant_id, bs = "re"),
            method = "REML",
            data = eyefitting_model_data)

@@

# Plot Predictions
eyefitting.gamm.plot <- eyefitting.grid.gamm %>%
  filter((parm_id %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(parm_id = factor(parm_id, levels = c("S", "F", "V", "N"))) %>%
  dplyr::rename(`Parameter Choice` = parm_id) %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data %>% mutate(parm_id = factor(parm_id, levels = c("S", "F", "V", "N"))) %>% dplyr::rename(`Parameter Choice` = parm_id), aes(x = x, y = residual_ols_drawn, group = plot_id, color = "OLS"), alpha = 0.2) +
  geom_line(data = eyefitting_model_data %>% mutate(parm_id = factor(parm_id, levels = c("S", "F", "V", "N"))) %>% dplyr::rename(`Parameter Choice` = parm_id), aes(x = x, y = residual_pca_drawn, group = plot_id, color = "PA"), alpha = 0.2) +
  geom_ribbon(aes(ymin = ols.lower, ymax = ols.upper, fill = "OLS"), color = NA, alpha = 0.4) +
  geom_line(aes(y = ols.pred, color = "OLS")) +
  geom_ribbon(aes(ymin = pca.lower, ymax = pca.upper, fill = "PA"), color = NA, alpha = 0.4) +
  geom_line(aes(y = pca.pred, color = "PA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, ncol = 1) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "right",
        plot.title   = element_text(size = 12, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line")
        ) +
  scale_y_continuous("Residual") +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("GAMM fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) 
eyefitting.gamm.plot
```

>>>>

**Rmd file containing code can be accessed [here](https://github.com/earobinson95/Eye-Fitting-Straight-Lines-in-the-Modern-Era/blob/main/analysis/you-draw-it-eyefitting-analysis.Rmd).**

@@

## GET DATA FROM DATABASE

```{r db}
library(RSQLite)
library(DBI)
library(digest)
library(lubridate)

last_date <- as.numeric(as_datetime("2021-05-18 18:00:00 CDT"))

# Connect to database
filename <- "you_draw_it_data.db"
sqlite.driver <- dbDriver("SQLite")
db_con <- dbConnect(sqlite.driver, dbname = filename)
# dbListTables(db_con)

# Import eyefitting parameter details
eyefitting_parameter_details <- dbReadTable(db_con,"eyefitting_parameter_details") %>%
  mutate(parm_id = as.factor(as.character(parm_id)))

# Import user data
users <- dbReadTable(db_con,"users") %>%
  filter(study_starttime > 1620152231, study_starttime < last_date) %>%
  select(-ip_address) %>%
  rowwise() %>%
  mutate(participant_id = as.factor(digest(paste(nick_name, study_starttime))))

# Import exponential prediction feedback data
eyefitting_feedback_data <- dbReadTable(db_con,"feedback") %>%
  filter(study_starttime > 1620152231, study_starttime < last_date) %>%
  filter(parm_id %in% c("S", "F", "V", "N")) %>%
  dplyr::rename(yols = y) %>%
  rowwise() %>%
  mutate(participant_id = as.factor(digest(paste(nick_name, study_starttime)))) %>%
  left_join(users, by = c("participant_id", "nick_name", "study_starttime")) %>%
  filter(recruitment != "I am the researcher") %>%
  rowwise() %>%
  mutate(plot_id = as.factor(digest(paste(nick_name, start_time, end_time)))) %>%
  select(study_starttime, participant_id, nick_name, plot_id, start_time, end_time, parm_id, x, yols, ydrawn, age, gender, academic_study, recruitment)

factorCols <- c("nick_name", "study_starttime", "age", "gender", "academic_study",
                "recruitment", "participant_id", "parm_id", "plot_id")
eyefitting_feedback_data[,factorCols] <- lapply(eyefitting_feedback_data[,factorCols], factor)
# write.csv(eyefitting_feedback_data, file = "eyefitting-feedback-data.csv", row.names = F, na = "")

# Import exponential prediction simulated data
eyefitting_simulated_data <- dbReadTable(db_con,"simulated_data") %>%
  filter(study_starttime > 1620152231, study_starttime < last_date) %>%
  filter(parm_id %in% c("S", "F", "V", "N")) %>%
  rowwise() %>%
  mutate(participant_id = as.factor(digest(paste(nick_name, study_starttime)))) %>%
  left_join(users, by = c("participant_id", "nick_name", "study_starttime")) %>%
  filter(recruitment != "I am the researcher")

factorCols <- c("nick_name", "dataset", "study_starttime", "age", "gender", "academic_study",
                "recruitment", "participant_id", "parm_id")
eyefitting_simulated_data[,factorCols] <- lapply(eyefitting_simulated_data[,factorCols], factor)
# write.csv(eyefitting_simulated_data, file = "eyefitting-simulated-data.csv", row.names = F, na = "")

# Disconnect from database
dbDisconnect(db_con)
length(unique(eyefitting_feedback_data$participant_id))
length(unique(eyefitting_feedback_data$plot_id))
```

@@

```{r, fig.width = 12, fig.height = 4}

eyefitting_parameter_details <- read_csv("eyefitting-parameter-details.csv")
eyefitting_parameter_details %>% knitr::kable()

@@

# write.csv(eyefitting_sim_data, "eyefitting-simulated-data-example.csv", row.names = F, na = "")

eyefitting_example_simplot <- eyefitting_sim_data %>%
  filter(dataset == "point_data") %>%
  # filter(dataset %in% c("F", "N", "S") | (x < 16 & x > 4)) %>%
  mutate(parm_id = factor(parm_id, levels = c("F", "S", "V", "N"))) %>%
  dplyr::rename(`Parameter Choice` = parm_id) %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(size = 1) +
  facet_wrap(~`Parameter Choice`, ncol = 4) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
  legend.position = "none",
  plot.title   = element_text(size = 12, hjust = 0),
  axis.text    = element_text(size = 12),
  axis.title   = element_text(size = 12),
  legend.title = element_text(size = 12),
  legend.text  = element_text(size = 12),
  # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
  # strip.background = element_rect(size = 0.5),
  legend.key.size = unit(1, "line")
) +
  scale_y_continuous(breaks = seq(-10, 20, 5))
eyefitting_example_simplot
```

@@

# save data to csv - already run once and pushed to GitHub.
# write.csv(eyefitting_model_data, file = "../data/eyefitting-model-data.csv", row.names = F, na = "")
```

@@

De-identified participant data collected in the study and used for analyses are available to be viewed and downloaded from GitHub at [here](https://github.com/earobinson95/Eye-Fitting-Straight-Lines-in-the-Modern-Era/raw/main/data).

```{r data}
eyefitting_model_data  <- read_csv("https://github.com/earobinson95/Eye-Fitting-Straight-Lines-in-the-Modern-Era/raw/main/data/eyefitting-model-data.csv")

@@

```{r, fig.height = 4, fig.width = 4}
# Example
eyefitting_model_data %>%
  filter(participant_id == "60f510d6896cddc508ea6d3de4b58234", parm_id == "F") %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = yols, color = "OLS", linetype = "OLS")) +
  geom_line(aes(y = ypca, color = "PA", linetype = "PA")) +
  geom_line(aes(y = ydrawn, color = "Drawn", linetype = "Drawn")) +
  geom_point(data = eyefitting_simulated_data %>%
               filter(dataset == "point_data", participant_id == "60f510d6896cddc508ea6d3de4b58234", parm_id == "F"),
             aes(x = x, y = y)) +
  facet_wrap(~ parm_id) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_y_continuous("y") +
  scale_color_manual("", values = c("black", "steelblue", "orange2")) +
  scale_linetype_manual("", values = c("dashed", "solid", "solid"))
```

@@

```{r y-ols-plot, fig.width = 12, fig.height = 4}

eyefitting_model_data %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = ydrawn, group = plot_id), alpha = 0.5, color = "steelblue") +
  geom_line(alpha = 0.2, aes(y = yols, group = participant_id)) +
  facet_wrap(~ parm_id, ncol = 4) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_x_continuous(limits = c(0, 20)) +
  ggtitle("OLS")
```

@@

```{r y-pca-plot, fig.width = 12, fig.height = 4}
eyefitting_model_data %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = ydrawn, group = plot_id), alpha = 0.5, color = "steelblue") +
  geom_line(alpha = 0.2, aes(y = ypca, group = participant_id)) +
  facet_wrap(~ parm_id, ncol = 4) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_x_continuous(limits = c(0, 20)) +
  ggtitle("PA")
```

@@

```{r residual-plot, fig.width = 12, fig.height = 4}
eyefitting_model_data %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = residual_ols_drawn, group = plot_id, color = "OLS"), alpha = 0.3) +
  geom_line(aes(y = residual_pca_drawn, group = plot_id, color = "PA"), alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~parm_id, ncol = 4) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_x_continuous(limits = c(0, 20)) +
  scale_color_manual("Estimate", values = c("steelblue", "orange"))
```

@@

+ parm_ID (F, S, V, N)
+ x (0, 20)

@@

```{r lmer, fig.width = 12, fig.height = 4}
# OLS
eyefitting.ols.lmer <- lmer(residual_ols_drawn ~ -1 + parm_id + x:parm_id + (1|participant_id),
                       data = eyefitting_model_data)
anova(eyefitting.ols.lmer)

@@

# Plot Predictions
eyefitting.lmer.plot <- eyefitting.preds.lmer %>%
  filter((parm_id %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(parm_id = factor(parm_id, levels = c("F", "S", "V", "N"))) %>%
  dplyr::rename(`Parameter Choice` = parm_id) %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data %>% mutate(parm_id = factor(parm_id, levels = c("F", "S", "V", "N"))) %>% dplyr::rename(`Parameter Choice` = parm_id), 
            aes(x = x, y = residual_ols_drawn, group = plot_id, color = "OLS"), alpha = 0.2) +
  geom_line(data = eyefitting_model_data %>% mutate(parm_id = factor(parm_id, levels = c("F", "S", "V", "N"))) %>% dplyr::rename(`Parameter Choice` = parm_id), 
            aes(x = x, y = residual_pca_drawn, group = plot_id, color = "PCA"), alpha = 0.2) +
  geom_ribbon(aes(ymin = asymp.LCL.ols, ymax = asymp.UCL.ols, fill = "OLS"), color = NA, alpha = 0.4) +
  geom_line(aes(y = emmean.ols, color = "OLS")) +
  geom_ribbon(aes(ymin = asymp.LCL.pca, ymax = asymp.UCL.pca, fill = "PA"), color = NA, alpha = 0.4) +
  geom_line(aes(y = emmean.pca, color = "PCA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, ncol = 4) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "right",
        plot.title   = element_text(size = 12, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line")
        ) +
  scale_y_continuous("Residual") +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("LMER fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PA"))

@@

```{r gamm, fig.width = 12, fig.height = 4}
# OLS
eyefitting.ols.gamm <- bam(residual_ols_drawn ~ parm_id + s(x, by = parm_id) + 
                             s(participant_id, bs = "re") +
                             s(x,participant_id, bs = "re"),
            method = "REML",
            data = eyefitting_model_data)

@@

# Plot Predictions
eyefitting.gamm.plot <- eyefitting.grid.gamm %>%
  filter((parm_id %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(parm_id = factor(parm_id, levels = c("F", "S", "V", "N"))) %>%
  dplyr::rename(`Parameter Choice` = parm_id) %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data %>% mutate(parm_id = factor(parm_id, levels = c("F", "S", "V", "N"))) %>% dplyr::rename(`Parameter Choice` = parm_id), aes(x = x, y = residual_ols_drawn, group = plot_id, color = "OLS"), alpha = 0.2) +
  geom_line(data = eyefitting_model_data %>% mutate(parm_id = factor(parm_id, levels = c("F", "S", "V", "N"))) %>% dplyr::rename(`Parameter Choice` = parm_id), aes(x = x, y = residual_pca_drawn, group = plot_id, color = "PA"), alpha = 0.2) +
  geom_ribbon(aes(ymin = ols.lower, ymax = ols.upper, fill = "OLS"), color = NA, alpha = 0.4) +
  geom_line(aes(y = ols.pred, color = "OLS")) +
  geom_ribbon(aes(ymin = pca.lower, ymax = pca.upper, fill = "PA"), color = NA, alpha = 0.4) +
  geom_line(aes(y = pca.pred, color = "PA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, ncol = 4) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "right",
        plot.title   = element_text(size = 12, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line")
        ) +
  scale_y_continuous("Residual") +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("GAMM fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) 
eyefitting.gamm.plot
```
